---
title: "All years"
output: html_document
params:
  filename: NA
---

```{r setup, include=FALSE}

library(tidyverse)
library(readxl)
library(DT)

knitr::opts_chunk$set(echo = FALSE)

all_data <- map(2 : 7, ~ rbind(read_excel(params$filename, 
                                          sheet = ., skip = 1,
                                          n_max = 31)))

names(all_data) = paste0("Y", 1:6)

list2env(all_data, envir = .GlobalEnv)

process_data <- function(input_data, top_section_variables, middle_section_variables, bottom_section_variables){
  
  input_data <- input_data %>%
    mutate_all(~ replace_na(.x, 0))
  
  input_data <- input_data %>% 
    rowwise() %>% 
    mutate(top = sum(across(all_of(top_section_variables))) / length(top_section_variables) / 3,
           middle = sum(across(all_of(middle_section_variables))) / length(middle_section_variables) / 3,
           bottom = sum(across(all_of(bottom_section_variables))) / length(bottom_section_variables) / 3 ,
           all = (sum(across(all_of(top_section_variables))) +
             sum(across(all_of(middle_section_variables))) +
             sum(across(all_of(bottom_section_variables)))) / (length(top_section_variables) + length(middle_section_variables) + length(bottom_section_variables)) / 3)
}

```

## R Markdown

* Below- if it doesn't do any of this
* Commencing- If a pupil gets at more than 50% of the total for the sheet they will be awarded at least commencing for that year group
* Developing - If a pupil scores 60% or more of the total score they will be awarded at least developing
* Secure - 100% from top section, and >50% from middle section inc middle green at 3
* Advanced - 100% from top section, >75% from middle section and >50% from bottom section
* Greater depth - 100% from top section >75% from middle section, >50% from bottom section inc green statement at 3

Output- 

* "Year 1 commencing", "Year 1 secure"...

```{r, results = "asis"}

if(nrow(Y1) > 0){
  
  cat("## Year 1")
}

```


```{r, year1}

if(nrow(Y1) > 0){
  
  names(Y1)[1] <- "child"

top_section_variables <- c(
  "I can name the letters of the alphabet in order", 
  "I can write simple sentences spoken by my teacher", 
  "I can talk about what I am going to write", 
  "I can compose my sentence aloud before I write it", 
  "I understand how words combine to make sentences", 
  "I can write my ideas down, sentence by sentence to write a short narrative (3+ sentences)", 
  "I can spell words using the phonemes I know", 
  "I can spell some tricky roof words correctly in my writing (see list)", 
  "I usually put a capital letter at the beginning and a full stop at the end of my sentence", 
  "I use finger spaces to separate words", 
  "I can check my writing makes sense by reading it through", 
  "I can begin to write lower letters correctly")

middle_section_variables <- c(
  "I can use letter names to distinguish between different spellings of the same sound", 
  "I can spell many roof words and all days of he week correctly", 
  "I can add s and es to make a plural", 
  "I can use the suffixes ing and ed", 
  "I can use a capital letter for names and for I", 
  "I can sit correctly at a table and hold my pencil correctly", 
  "I can form capital letters and 0-9", 
  "I can discuss what I have written and read it aloud audibly")

bottom_section_variables <- c(
  "I can write a short narrative (5+ sentences) that follows a suitable structure", 
  "I can begin to use ? And ! In my writing", 
  "I can spell most roof words correctly (up to card 21)", 
  "I can use the prefix un- in my writing", 
  "I can use the suffixes er and est in my writing", 
  "I can use and to join 2 sentences")

# fix missing variables

Y1 <- process_data(Y1, top_section_variables, middle_section_variables, bottom_section_variables)
# calculate scores

# commencing - 38 from anywhere 50%
# developing - 45 from anywhere 60%
# secure - 33 from top section, and 13 from middle section
# advanced - 33 from top section, 18 from middle section and 6 from bottom section
# greater depth - 33 from top section 18 from middle section, 9 from bottom section inc green statement

Y1 <- Y1 %>% 
  mutate(stage = case_when(
    top == 1 & middle >= .75 & bottom >= .5 & `I can write a short narrative (5+ sentences) that follows a suitable structure` == 3 ~ "f_deep",
    top == 1 & middle >= .75 & bottom >= .5 ~ "e_advanced",
    top == 1 & middle >= .5  ~ "d_secure",
    all >= .6 ~ "c_developing",
    all >= .5 ~ "b_commencing",
    TRUE ~ "a_none"
  ))

Y1 %>% 
  select(child, top, middle, bottom, 
         `I can write a short narrative (5+ sentences) that follows a suitable structure`, 
         all, stage) %>% 
  set_names(c("Name", "Ready to progress", "Age-related expectations",
              "Above age-related", "Greater depth required",
              "Total score", "Assessment level")) %>% 
  datatable(rownames = FALSE, extensions = 'Buttons', options = list(
    pageLength = 100,
    dom = 'Bfrtip',
    buttons = c('copy', 'csv', 'excel', 'pdf', 'print')
  )) %>% 
  formatRound(c("Ready to progress", "Age-related expectations",
                "Above age-related", "Greater depth required", "Total score"), 2)
}

```

```{r, results = "asis"}

if(nrow(Y2) > 0){
  
  cat("## Year 2")
}

```

```{r, year2}

if(nrow(Y2) > 0){
  
  names(Y2)[1] <- "child"
  
  top_section_variables2 <- c("Write simple, coherent narratives about personal experiences and those of others", 
                              "Write about real events, recording these simply and clearly.", 
                              "Demarcate most sentences in their writing with capital letters and full stops", 
                              "Use present and past tense mostly correctly and consistently.", 
                              "Use co-ordination (e.g. or / and / but) to join clauses.", 
                              "I can use a range of suffixes appropriately: -s, -es, -er, -ed, -ing", 
                              "segment spoken words into phonems and represent these with graphemes (some correct spellings, some phonetically plausible)")
  
  middle_section_variables2 <- c("Use some subordination (e.g. when / if / that / because) to join clauses.", 
                                 "Use question marks correctly when appropriate", 
                                 "segment spoken words into phonems and represent these with graphemes (many correct spellings, some phonetically plausible)", 
                                 "Spell many common exception words (Y2 spelling list – up to card 30)", 
                                 "form capital letters and digits of he correct size, orientation and relationship to each other and lowercase", 
                                 "Use spacing between words that reflects the size of the letters.")
  
  bottom_section_variables2 <- c("Write effectively and coherently for different purposes.", 
                                 "Draw on their reading to inform the vocabulary and grammar of their writing.", 
                                 "Make simple additions, revisions and proof-reading corrections to their own writing.", 
                                 "Use punctuation taught at key stage 1 mostly correctly", 
                                 "Spell most common exception words.", 
                                 "add suffixes to spell most words  correctly in their writing (ment, ness, ful, ly)", 
                                 "Use the diagonal and horizontal strokes needed to join some letters.")
  
  # fix missing variables
  
  Y2 <- process_data(Y2, top_section_variables2, middle_section_variables2, bottom_section_variables2)
  
  Y2 <- Y2 %>% 
    mutate(stage = case_when(
      top == 1 & middle >= .75 & bottom >= .5 & `Write effectively and coherently for different purposes.` == 3 ~ "f_deep",
      top == 1 & middle >= .75 & bottom >= .5 ~ "e_advanced",
      top == 1 & middle >= .5  ~ "d_secure",
      all >= .6 ~ "c_developing",
      all >= .5 ~ "b_commencing",
      TRUE ~ "a_none"
    ))
  
  Y2 %>% 
    select(child, top, middle, bottom, 
           `Write effectively and coherently for different purposes.`, 
           all, stage) %>% 
    set_names(c("Name", "Ready to progress", "Age-related expectations",
                "Above age-related", "Greater depth required",
                "Total score", "Assessment level")) %>% 
    datatable(rownames = FALSE, extensions = 'Buttons', options = list(
      pageLength = 100,
      dom = 'Bfrtip',
      buttons = c('copy', 'csv', 'excel', 'pdf', 'print')
    )) %>% 
    formatRound(c("Ready to progress", "Age-related expectations",
                  "Above age-related", "Greater depth required", "Total score"), 2)
  
}

```

```{r, results = "asis"}

if(nrow(Y3) > 0){
  
  cat("## Year 3")
}

```


```{r, year3}

if(nrow(Y3) > 0){
  
  names(Y3)[1] <- "child"
  
  top_section_variables3 <- c("Write for different purposes using text and grammar features expected in narrative and factual pieces", 
                              "Use a range of sentence types which are mostly punctuated correctly with capital letter, full stop, exclamation mark or question mark", 
                              "Use co-ordinating conjunctions to combine 2 main clauses - FANBOYS.", 
                              "Use some subordination (e.g. when / if / that / because) to join clauses", 
                              "Proof-read my work for simple errors.", "Use the horizontal and diagonal strokes needed to join some letters.", 
                              "Spell a few homophones and most Y2 common exception words accurately (roof words up to card 30)", 
                              "Use simple organisational devices and organise paragraphs around a theme.", 
                              "Use adverbs of time, place and cause e.g. then, next, soon, therefore")
  
  middle_section_variables3 <- c("Write coherently and effectively for different purposes including narrative and factual pieces and my wriing contains interesting details", 
                                 "Use correct punctuation in all my writing, e.g. commas in a list, apostrophes for contraction and possession, beginning to use inverted commas for direct speech **", 
                                 "Use prepositions to express time, place and cause e.g. before, after, during, in, because, of.", 
                                 "Use conjunctions to express time and cause e.g. when, before, after, while, so and because", 
                                 "Spell some homophones and many Y3 common exception words accurately.", 
                                 "Use a range of pre-fixes and suffixes e.g.: dis, mis, un, de, and –ly, -ment, -ness, -ful, -less.", 
                                 "Proof-read my work for punctuation and spelling errors.", "Use diagonal and horizontal strokes to join letters and my writing is spaced sufficiently to aid legibility"
  )
  
  bottom_section_variables3 <- c("Write coherently and effectively in a range of genres. My pieces include a variety of genre specific vocab which adds interest to the reader", 
                                 "Successfully create settings, characters and plots in narrative writing ensuring it follows a sequence", 
                                 "Use the present perfect form of verbs instead of simple past tense.", 
                                 "Use a or an depending on if a word starts with a consonant or vowel.", 
                                 "Use inverted commas accurately e.g. The conductor shouted, “Sit down!”", 
                                 "Proof-read my work for simple grammatical errors and check that my writing makes sense.", 
                                 "Spell many homophones and common exception words from the Y3/4 list accurately", 
                                 "Write ascenders and descenders are equidistant and parallel."
  )
  
  # fix missing variables
  
  Y3 <- process_data(Y3, top_section_variables3, middle_section_variables3, bottom_section_variables3)
  # calculate scores
  
  # commencing - 38 from anywhere 50%
  # developing - 45 from anywhere 60%
  # secure - 33 from top section, and 13 from middle section
  # advanced - 33 from top section, 18 from middle section and 6 from bottom section
  # greater depth - 33 from top section 18 from middle section, 9 from bottom section inc green statement
  
  Y3 <- Y3 %>% 
    mutate(stage = case_when(
      top == 1 & middle >= .75 & bottom >= .5 & `Write coherently and effectively in a range of genres. My pieces include a variety of genre specific vocab which adds interest to the reader` == 3 &
        `Use correct punctuation in all my writing, e.g. commas in a list, apostrophes for contraction and possession, beginning to use inverted commas for direct speech **` == 3 ~ "f_deep",
      top == 1 & middle >= .75 & bottom >= .5 ~ "e_advanced",
      top == 1 & middle >= .5  ~ "d_secure",
      all >= .6 ~ "c_developing",
      all >= .5 ~ "b_commencing",
      TRUE ~ "a_none"
    ))
  
  Y3 %>% 
    select(child, top, middle, bottom, 
           `Write coherently and effectively in a range of genres. My pieces include a variety of genre specific vocab which adds interest to the reader`,
           `Use correct punctuation in all my writing, e.g. commas in a list, apostrophes for contraction and possession, beginning to use inverted commas for direct speech **`, 
           all, stage) %>% 
    set_names(c("Name", "Ready to progress", "Age-related expectations",
                "Above age-related", "Greater depth required", "Greater depth required_2", 
                "Total score", "Assessment level")) %>% 
    datatable(rownames = FALSE, extensions = 'Buttons', options = list(
      pageLength = 100,
      dom = 'Bfrtip',
      buttons = c('copy', 'csv', 'excel', 'pdf', 'print')
    )) %>% 
    formatRound(c("Ready to progress", "Age-related expectations",
                  "Above age-related", "Greater depth required", "Total score"), 2)
}

```

```{r, results = "asis"}

if(nrow(Y4) > 0){
  
  cat("## Year 4")
}

```


```{r, year4}

if(nrow(Y4) > 0){
  
  names(Y4)[1] <- "child"
  
  top_section_variables4 <- c("I can use fronted adverbials with commas", "I can use inverted commas and other punctuation to indicate direct speech", 
                              "I can extend my sentences to more than one clause using a wider range of conjunctions including: when, if, because and although", 
                              "I can expand my noun phrases by adding adjectives and preposition phrases to the noun", 
                              "I can consistently use the correct tense", "I can use nouns and pronouns to avoid repetition.", 
                              "I can use paragraphs to organise my ideas around a theme**", 
                              "I can make my writing engaging and fluent so that the reader wants to read on", 
                              "I can spell words using the rules taught in Year 2", "I can join my handwriting in a consistent and legible style")
  
  middle_section_variables4 <- c("I can tell the difference between a plural s and a possessive s", 
                                 "I use the correct verb inflection – we were, I did", "I can use a range of determiners", 
                                 "I can write narratives with a clear beginning, middle and end", 
                                 "I can use a range of organisational devices in non-fiction writing; e.g sub-headings and bullet points**", 
                                 "I can plan by discussing writing that is similar to the piece I want to write", 
                                 "I can develop descriptions of characters, settings and emotions**", 
                                 "I can use adjectives and adverbs when writing descriptions", 
                                 "I can proofread confidently and amend my own and others’ spellings, grammar and punctuation.", 
                                 "I can suggest improvements in my own and others’ writing so that there is a better effect on the reader.", 
                                 "I can spell words using the rules taught in Year 3 and 4")
  
  bottom_section_variables4 <- c("I can consistently use possessive apostrophes correctly", "I can use conjunctions and adverbs to express time and cause", 
                                 "I can use rich and varied vocabulary", "I can begin to include my opinions, interests and humour to engage the reader in a range of genres", 
                                 "I can spell all/ almost all of the words correctly from the Y3 and 4 statutory spelling list**")
  
  # fix missing variables
  
  Y4 <- process_data(Y4, top_section_variables4, middle_section_variables4, bottom_section_variables4)
  # calculate scores
  
  # commencing - 38 from anywhere 50%
  # developing - 45 from anywhere 60%
  # secure - 33 from top section, and 13 from middle section
  # advanced - 33 from top section, 18 from middle section and 6 from bottom section
  # greater depth - 33 from top section 18 from middle section, 9 from bottom section inc green statement
  
  Y4 <- Y4 %>% 
    mutate(stage = case_when(
      top == 1 & middle >= .75 & bottom >= .5 & `I can begin to include my opinions, interests and humour to engage the reader in a range of genres` == 3 ~ "f_deep",
      top == 1 & middle >= .75 & bottom >= .5 ~ "e_advanced",
      top == 1 & middle >= .5  ~ "d_secure",
      all >= .6 ~ "c_developing",
      all >= .5 ~ "b_commencing",
      TRUE ~ "a_none"
    ))
  
  Y4 %>% 
    select(child, top, middle, bottom, 
           `I can begin to include my opinions, interests and humour to engage the reader in a range of genres`,
           all, stage) %>% 
    set_names(c("Name", "Ready to progress", "Age-related expectations",
                "Above age-related", "Greater depth required", 
                "Total score", "Assessment level")) %>% 
    datatable(rownames = FALSE, extensions = 'Buttons', options = list(
      pageLength = 100,
      dom = 'Bfrtip',
      buttons = c('copy', 'csv', 'excel', 'pdf', 'print')
    )) %>% 
    formatRound(c("Ready to progress", "Age-related expectations",
                  "Above age-related", "Greater depth required", "Total score"), 2)
}

```


```{r, results = "asis"}

if(nrow(Y5) > 0){
  
  cat("## Year 5")
}

```

```{r, year5}

if(nrow(Y5) > 0){
  
  names(Y5)[1] <- "child"
  
  top_section_variables5 <- c("I can use a range of punctuation in all my writing; e.g apostrophes, inverted commas, comma, exclamation marks and question marks**", 
                              "I can write multi-clause sentences using commas to separate clauses", 
                              "I can write using the correct tense", "I can use paragraphs to organise my ideas in fiction and non-fiction**", 
                              "I can plan my writing with a clear structure, purpose and audience.", 
                              "I can write descriptions of settings, characters and atmosphere that use a range of devices**", 
                              "I can spell words using the rules taught in Year 5", "I can join my handwriting in a clear and legible style")
  
  middle_section_variables5 <- c("I can use expanded noun phrases.", "I can use adverbs.", 
                                 "I can use a variety of co-ordinating and subordinating conjunctions.", 
                                 "I can develop my ideas by drawing on prior reading and research.", 
                                 "I can write narratives with a clear opening, build up, problem, resolution and ending.", 
                                 "I can use a range of organisational and presentational devices in non-fiction writing, e.g. sub-headings and bullet points**", 
                                 "I can write with different levels of formality depending on the text type.", 
                                 "I can proofread confidently and amend my own and others’ spellings, grammar and punctuation and suggest improvements.",
                                 "I can publish my work to a high standard, including using ICT.", 
                                 "I can spell words using the rules taught in year three and four.")
  
  bottom_section_variables5 <- c("I can use brackets, dashes and commas to show parenthesis.", 
                                 "I can write relative clauses beginning with relative pronouns.", 
                                 "I can use modal verbs to show possibility", "I can use cohesive devices within paragraphs and across paragraphs", 
                                 "I can include my opinions, interests and humour in my writing", 
                                 "I can use techniques to develop the readers’ interest, e.g. building tension, metaphors, similes, rhetorical questions in a range of genres", 
                                 "I can use dialogue to show what a character is like and to show their actions.", 
                                 "I can spell most of the words correctly from the Y5 statutory spelling list")
  
  # fix missing variables
  
  Y5 <- process_data(Y5, top_section_variables5, middle_section_variables5, bottom_section_variables5)
  
  # commencing - 38 from anywhere 50%
  # developing - 45 from anywhere 60%
  # secure - 33 from top section, and 13 from middle section
  # advanced - 33 from top section, 18 from middle section and 6 from bottom section
  # greater depth - 33 from top section 18 from middle section, 9 from bottom section inc green statement
  
  Y5 <- Y5 %>% 
    mutate(stage = case_when(
      top == 1 & middle >= .75 & bottom >= .5 & `I can use techniques to develop the readers’ interest, e.g. building tension, metaphors, similes, rhetorical questions in a range of genres` == 3 ~ "f_deep",
      top == 1 & middle >= .75 & bottom >= .5 ~ "e_advanced",
      top == 1 & middle >= .5  ~ "d_secure",
      all >= .6 ~ "c_developing",
      all >= .5 ~ "b_commencing",
      TRUE ~ "a_none"
    ))
  
  Y5 %>% 
    select(child, top, middle, bottom, 
           `I can use techniques to develop the readers’ interest, e.g. building tension, metaphors, similes, rhetorical questions in a range of genres`, 
           all, stage) %>% 
    set_names(c("Name", "Ready to progress", "Age-related expectations",
                "Above age-related", "Greater depth required",
                "Total score", "Assessment level")) %>% 
    datatable(rownames = FALSE, extensions = 'Buttons', options = list(
      pageLength = 100,
      dom = 'Bfrtip',
      buttons = c('copy', 'csv', 'excel', 'pdf', 'print')
    )) %>% 
    formatRound(c("Ready to progress", "Age-related expectations",
                  "Above age-related", "Greater depth required", "Total score"), 2)
}

```

```{r, results = "asis"}

if(nrow(Y6) > 0){
  
  cat("## Year 6")
}

```

```{r, year6}

if(nrow(Y6) > 0){
  
  names(Y6)[1] <- "child"
  
  top_section_variables6 <- c("Use correct punctuation in all my writing, e.g. commas in a list, apostrophes for contraction and possession, beginning to use inverted commas for direct speech **", 
                              "I can use paragraphs to organise my ideas around a theme**", 
                              "I can use a range of organisational devices in non-fiction writing; e.g sub-headings and bullet points**", 
                              "I can develop descriptions of characters, settings and emotions**", 
                              "I can spell all/ almost all of the words correctly from the Y3 and 4 statutory spelling list**", 
                              "I can use paragraphs to organise my ideas in fiction and non-fiction**", 
                              "I can write descriptions of settings, characters and atmosphere that use a range of devices**")
  
  middle_section_variables6 <- c("I can select vocabulary and grammatical structures that reflect what the writing requires, doing this mostly appropriately", 
                                 "I can use verb tenses consistently and correctly throughout my writing", 
                                 "I can use the range of punctuation taught at key stage 2 mostly correctly", 
                                 "I can write effectively for a range of purposes and audiences, selecting language that shows good awareness of the reader", 
                                 "I can use a range of devices to build cohesion within and across paragraphs", 
                                 "I can describe settings, characters and atmosphere in narratives", 
                                 "I can integrate dialogue in narratives to convey character and advance the action", 
                                 "I can spell correctly most words from the year 5/ year 6 spelling list", 
                                 "I can maintain legibility in joined handwriting when writing at speed")
  
  bottom_section_variables6 <- c("I can use the range of punctuation taught at key stage 2 and, when necessary, use the punctuation precisely to enhance meaning and avoid ambiguity", 
                                 "I can write effectively for a range of purposes and audiences, selecting the appropriate form and drawing independently on what I have read as models for my own writing", 
                                 "I can exercise an assured and conscious control over levels of formality, particularly through manipulating grammar and vocabulary to achieve this", 
                                 "I can distinguish between the language of speech and writing and choose the appropriate register")
  
  # fix missing variables
  
  Y6 <- process_data(Y6, top_section_variables6, middle_section_variables6, bottom_section_variables6)
  
  # commencing - 38 from anywhere 50%
  # developing - 45 from anywhere 60%
  # secure - 33 from top section, and 13 from middle section
  # advanced - 33 from top section, 18 from middle section and 6 from bottom section
  # greater depth - 33 from top section 18 from middle section, 9 from bottom section inc green statement
  
  Y6 <- Y6 %>% 
    mutate(stage = case_when(
      top == 1 & middle >= .75 & bottom >= .5 & `I can write effectively for a range of purposes and audiences, selecting the appropriate form and drawing independently on what I have read as models for my own writing` == 3 ~ "f_deep",
      top == 1 & middle >= .75 & bottom >= .5 ~ "e_advanced",
      top == 1 & middle >= .5  ~ "d_secure",
      all >= .6 ~ "c_developing",
      all >= .5 ~ "b_commencing",
      TRUE ~ "a_none"
    ))
  
  Y6 %>% 
    select(child, top, middle, bottom, 
           `I can write effectively for a range of purposes and audiences, selecting the appropriate form and drawing independently on what I have read as models for my own writing`, 
           all, stage) %>% 
    set_names(c("Name", "Ready to progress", "Age-related expectations",
                "Above age-related", "Greater depth required",
                "Total score", "Assessment level")) %>% 
    datatable(rownames = FALSE, extensions = 'Buttons', options = list(
      pageLength = 100,
      dom = 'Bfrtip',
      buttons = c('copy', 'csv', 'excel', 'pdf', 'print')
    )) %>% 
    formatRound(c("Ready to progress", "Age-related expectations",
                  "Above age-related", "Greater depth required", "Total score"), 2)
}

```

